E Coli Project Paper Incredibly Rough Draft

### 2013-09-30 paper/paper.tex

\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre]{revtex4-1}

\usepackage{graphicx}% Include figure files
\usepackage{color}

\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\blue}[1]{{\bf \color{blue} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}

@@

@@

\author{Jeff B. Schulte}
\affiliation{Department of Physics, Oregon State University}
\author{Rene WHAT. Zeto}
\affiliation{Department of Physics, Oregon State University}
\author{David Roundy}
\affiliation{Department of Physics, Oregon State University}

\begin{abstract}
  The dynamics of the Min-protein system help Escherichia coli
  bacteria regulate the process of cell division by identifying the
  center of the cell.  We model the Min-protein system in bacteria
  that have been forced into unusual flattened shapes, as have
  recently been experimentally observed.  We find that although the
  presence of Min oscillations is robust in a wide variety of cellular
  configurations, the location of the peaks is strongly affected by
  the cellular shape.  In some cases no periodic oscillations are
  observed.  In particular, we find that cellular shapes observed
  experimentally to present irregular oscillations do so in the
  theoretical model, consistently \fixme{or inconsistently?} with
  experiment.  \fixme{In agreement with previous theoretical and
    experimental results, we observe ``rotating'' behavior in certain
    shapes having three corners.}
\end{abstract}

\maketitle

\section{Introduction}
It is vital that during the process of bacterial cell division a cell
avoid minicelling, or splitting into daughter cells with lopsided
volumes.  During this process a long FtsZ polymer chain develops on
the cell wall in the center region of the cell that dictates the plane
of division\cite{adams2009bacterial}
\cite{lutkenhaus2007assembly}. Previous experimental studies have
shown that the MinC protein, known to inhibit the FtZ
polymer\cite{shen2010examination}, exhibits a pole to pole oscillatory
behavoir in conjunction with the MinD and MinE proteins, while the
oscillating MinE tends to be more localized in the center of the cell
\cite{hu1999topological} \cite{fu2001mine} \cite{shapiro2009and}
\cite{yu1999ftsz} \cite{raskin1999rapid}The MinC protein will then have a higher time
averaged concentration in the cell poles as opposed to the center
region of the cell, aiding in prohibiting the FtZ from developing in the wrong
region.

@@

Figure \ref{ave-per-area-plot-pill} shows the ave protein per unit
area that is wall attached in the middle, right, and left sections of
the pill shaped cell.  This is an important test of the central idea
behind the utility of the Min protein system.  The system is meant to
place Min proteins on the walls in heavier concentrations in the caps
of the cell and lower concentrations in the middle section, so that
the FtZ polymer may build up along the center walls.  We can see from
the plot of the 3$\mu m$ by 1$\mu m$ cell that the MinD proteins reach
concentration peaks in the caps that are roughly three times higher
(for the MinD-ATP proteins) and roughly 1.5 times higher (for the
MinD-MinE-ATP) than peaks in the center.  Integrals of these plotted
values show that the time averaged concentrations are likewise higher
in the caps than in the center.  \fixme{am computing these integrals
  now in ave\_plot.py but need to print out into maybe a bar chart}

@@

\fixme{Note to make sure I'm interpreting the iscosolese triangle data
  correctly: the triangle sides are 5,5,3, and the higher density
  starts in the corner that is opposite the short side.  This corner
  is covered by the right section.  I've verified this by looking at
  the protein\_microscopy and drawing a little picture, which is on my
  desk...if you dare look for something on my desk!!!}  The iscosolese
triangle starts its simulation with a higher concentration of protein
in the section that is opposite the shorter side.  It quickly relaxes
into a stable oscillation pattern.  This pattern shows that right
corner section regularly has higher spikes in total protein than the
other sections, and that its spikes appear one pi removed from the
spikes for the other two sections.  This a similar behavoir to that
which is seen in the equilateral triangle, differening in that it
converges more quickly towards this behavoir and that it shows a
greater exageration in the height of protein buildup in the right most
section.

>>>>

\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre]{revtex4-1}
\usepackage{graphicx}% Include figure files
\usepackage{animate}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{color}
\usepackage{breqn}

\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\blue}[1]{{\bf \color{blue} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}
\newcommand{\rr}{\textbf{r}}
\newcommand{\refnote}{\red{[ref]}}

@@

%\newcommand{\derivation}[1]{#1} % Use this to show all derivations in detail
\newcommand{\derivation}[1]{} % Use this for nice pegagogical paper...

% needsworklater is used to annotate bits that need work, but that we
% can postpone for a while.
\newcommand{\needsworklater}[1]{\emph{[#1]}}
% needsworknow is intended to prioritize stuff that needs fixing.
\newcommand{\needsworknow}[1]{\textcolor{red}{[\emph{#1}]}}

@@

%\pacs{61.20.Ne, 61.20.Gy, 61.20.Ja}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  This paper is about science.
\end{abstract}
\section{Introduction}
It is vital that during the process of bacterial cell division a cell
avoid minicelling, or splitting into daughter cells with lopsided
volumes.  During this process a long FtsZ polymer chain develops on
the cell wall in the center region of the cell that dictates the plane
of division\cite{adams2009bacterial}
\cite{lutkenhaus2007assembly}. Previous experimental studies have
shown that the MinC protein, known to inhibit the FtZ
polymer\cite{shen2010examination}, exhibits a pole to pole oscillatory
behavoir in conjunction with the MinD and MinE proteins, while the
oscillating MinE tends to be more localized in the center of the cell
\cite{hu1999topological} \cite{fu2001mine} \cite{shapiro2009and}
\cite{yu1999ftsz} \cite{raskin1999rapid}The MinC protein will then have a higher time
averaged concentration in the cell poles as opposed to the center
region of the cell, aiding in prohibiting the FtZ from developing in the wrong
region.

@@

Figure \ref{ave-per-area-plot-pill} shows the ave protein per unit
area that is wall attached in the middle, right, and left sections of
the pill shaped cell.  This is an important test of the central idea
behind the utility of the Min protein system.  The system is meant to
place Min proteins on the walls in heavier concentrations in the caps
of the cell and lower concentrations in the middle section, so that
the FtZ polymer may build up along the center walls.  We can see from
the plot of the 3$\mu m$ by 1$\mu m$ cell that the MinD proteins reach
concentration peaks in the caps that are roughly three times higher
(for the MinD-ATP proteins) and roughly 1.5 times higher (for the
MinD-MinE-ATP) than peaks in the center.  Integrals of these plotted
values show that the time averaged concentrations are likewise higher
in the caps than in the center.  \fixme{am computing these integrals
  now in ave_plot.py but need to print out into maybe a bar chart}

@@

\fixme{Note to make sure I'm interpreting the iscosolese triangle data
  correctly: the triangle sides are 5,5,3, and the higher density
  starts in the corner that is opposite the short side.  This corner
  is covered by the right section.  I've verified this by looking at
  the protein_microscopy and drawing a little picture, which is on my
  desk...if you dare look for something on my desk!!!}  The iscosolese
triangle starts its simulation with a higher concentration of protein
in the section that is opposite the shorter side.  It quickly relaxes
into a stable oscillation pattern.  This pattern shows that right
corner section regularly has higher spikes in total protein than the
other sections, and that its spikes appear one pi removed from the
spikes for the other two sections.  This a similar behavoir to that
which is seen in the equilateral triangle, differening in that it
converges more quickly towards this behavoir and that it shows a
greater exageration in the height of protein buildup in the right most
section.
